---
title: "Data Manipulation with dplyr"
author: "Shun Xie"
date: "2022-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
  )
```

# Cleaning

### Select columns 

Once you've import the data, you're going to need to do some cleaning up. 

```{r}
library(tidyverse)
options(tibble.print_min=3)

litters_data = read_csv('data/FAS_litters.csv')
litters_data = janitor::clean_names(litters_data)

```


```{r}
pups_datat =read_csv('data/FAS_pups.csv')
pups_datat = janitor::clean_names(pups_datat)

```

```{r}
select(litters_data,group,litter_number, gd0_weight,pups_born_alive)
```




```{r}
#select a range of variables 
select(litters_data, group:gd_of_birth)
```
You can also specify columns of you'd like to remove

```{r}
select(litters_data,-pups_survive)

select(litters_data,-pups_survive,-group)

```

You can rename variables as part of this process
```{r}
select(litters_data,GROUP=group, LiTtEr_NuMbEr=litter_number)
```

only rename the variable and keep all of the variables. 
```{r}
rename(litters_data,GROUP=group, LiTtEr_NuMbEr=litter_number)
```

start_with() ends_with() and contains()

```{r}
#View(litters_data)

select(litters_data,starts_with("gd"))
select(litters_data,ends_with("th"))

```


Change the ordering, take litters_data and litter_number to the front. Everything() will grab the other variables.
```{r}
select(litters_data, litter_number,pups_survive,everything())
```

### Select Rows (Filter)

">,>=,<,<=,== and !="

'%in%', 'is.na()'

group %in% c("Con7","Con8") (useful for character variable) 

* drop_na(litters_data) remove any row with a missing value

* drop_na(litters_data, wt_increase) will remove rows for which 'wt_increase' is missing



### 'mutate' (create variable)
```{r}
#create a new variable called wt_gain

#litter_data2 = 
mutate(litters_data,wt_gain = gd18_weight - gd0_weight, group = str_to_lower(group),
#wt_gain_kg=wt_gain*2.2
)

```

### arrange

arrange the rows in your data according to the values in one or more columns. 

```{r}
#sort first on group, then on pups_born_alive
head(arrange(litters_data,group, pups_born_alive),10)


#decending order
head(arrange(litters_data,desc(group), pups_born_alive),10)
```


```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")

```



### Piping %>%
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv",col_types = 'ccddiiii') %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>%
  drop_na(wt_gain)
  
  
litters_data
```

```{r}
#can use nontidyverse function as long as specified the starting?

litters_data %>%
  lm(wt_gain=pups_born_alive,data=.) %>%
  broom::tidy()
```






